Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-03-11 response_to_review_SOLA-D-20-00015_v1.txt

Dear Editor,

Thank you very much for your email with the report on the first version of our manuscript ID SOLA-D-20-00015, titled "Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling".

We deeply acknowledge the reviewer, whose report reflects her/his very careful and interested read. All queries very relevant, with useful and constructive criticisms and suggestions. Below, we copy/paste the full review. Following each query (Q) we include a corresponding answer (A), commenting the changes in the manuscript, in which all edited/new text is highlighted in red color font.

We addressed all points, as a result of which we believe the manuscript has improved in both content and clarity. In particular, a major improvement in this new version of the draft relates to Q5 below.

@@

-----------------------------------------------------------------------------------------
Q4. Table 2 summarises the differences of the median values of the DEMT and MHD model. This is OK. However, there is an opportunity here to present a more detailed comparison, including in particular the mean (or median) absolute deviation (MAD) of values between DEMT and MHD. The MAD is a more stringent, voxel-by-voxel test of the comparison, which cannot be seen from the histogram comparisons (e.g. figure 12/13) nor the median values of Table 2.

A4. When selecting field lines for analysis of both the DEMT and the AWSoM models, the classification of types of legs introduced in section 2.3 is applied to the DEMT reconstruction and the AWSoM model separately. As a result, as shown by comparing the middle panels of Figure 3 and the bottom panels of Figure 11, and as discussed in the corresponding texts of the manuscript, the selection of field lines for DEMT is not the same as the selection of field lines for the AWSoM model. In comparing those Figures each color indicates comparable regions between the two analysis. It is so that Figures 12 and 13, and Table 2, which show DEMT-versus-AWSoM comparison for each rotation, are not done line-by-line but rather region-by-region (orange, red and cyan). The suggestion of calculating the MAD would require selecting the same field lines for both models, which is beyond the scope of the present comparative analysis. Informing the MAD is a good idea, being a robust quantitative measure of how the comparison between both models spreads. We will certainly implement such an analysis in a next effort. In the present draft, we trust the current comparison (Figures 12 and 13, and Table 2) provide a pretty informative summary of how both models compare. Let us know what you think.

@@

-----------------------------------------------------------------------------------------
Q5. You discuss the less reliable results from AIA in coronal holes. The very recent paper by Saqri et al (Solar Physics 295, 2020) is relevant here, and should be referenced.

A5. Thanks for the reference, which is certainly relevant and now mentioned in the paper, as discussed below. We are not sure stray light contamination is actually the fundamental reason of the reported issue, which is now solved, as described next.

The discussion section of v1 of the draf includes this sentence (now deleted): "The source of this distinct performance of DEMT when applied to EUVI or AIA data is related to the TRFs of the respective filter sets. We are currently implementing an improved LDEM determination algorithm to deal with this issue that will be used in a follow up article." While waiting for your report and preparing V2 of the draft, we successfully implemented the improved version of the code, as a result of which this problem has been solved. The issue was related to the algorithm that solves the optimization problem posed in each voxel to find the LDEM. When based on AIA data, the previous algorithm tended to converge to a non-optimal (local minimum of the objective function) in specific voxels located in the CHs. We addressed the problem by exploring multiple starting points when solving the optimization problem in cases where the initial solution was not good enough. In this draft, the improved algorithm affects results concerning type III legs for CR2208. As a result, the following Figures have been updated: 2, 3, 4, 5, 10, 13 and 14. You will notice all these Figures are just like the original ones, except in the CHs of CR-2208, where the previous issue has been solved. The text throughout the manuscript was edited accordingly, in the following pages of V2 of the draft:
 - Related text in Pgs. 12-15 (some of the red text in these pages relates to Q4 above).
 - Numerical values concerning legs of type-III for CR-2208 in Tables 1 and 2.
 - Most of the third paragraph of the Conclusion section in V1 was devoted to this issue, so it was eliminated. The remaining text is now the end of the second paragraph of that section (highlighted in red color).

@@

-----------------------------------------------------------------------------------------
Q6. The abstract can be shortened somewhat, reducing the number of introductory sentences and with a little less emphasis on the method and future improvements, and more on the science results. In particular: the results of the energy balance analysis and the estimate of energy flux at the coronal base, which are important.

A6. We fully agree. We edited the Abstract, reducing the number of introductory lines from 12 down to 5, and greatly expanding the details in the results, and still keeping the full abstract 2 lines shorter than in V1. The presentation of results is better organized. We first summarize purely DEMT results. For this we were careful to mention which findings were first reported in previous DEMT studies. We still believe these are worth highlighting here as these were now found using the latest implementation of DEMT. Then we summarize how the results of the latest implementation of AWSoM compare with DEMT, highlighting the shortcomings as well.

@@

-----------------------------------------------------------------------------------------
Q8. The introduction does not have many references to works beyond DEMT and AWSoM. The 2nd paragraph of the introduction offers a chance to remedy this, where you discuss the quiet-Sun corona. I know there are very few works that address the quiet corona in the context of DEM analysis, but you can reference https://www.aanda.org/articles/aa/full_html/2014/04/aa23054-13/aa23054-13.html. You can also mention https://iopscience.iop.org/article/10.1088/0004-637X/795/2/111 here (although you do reference this paper in more detail in the discussion). Another relevant paper which can be referenced, with DEM analysis of a large amount of quiet coronal data over long time periods, is https://advances.sciencemag.org/content/3/7/e1602056 .

A8. We agree it is relevant to expand the references in the Introduction to other DEM studies of the quiet-Sun. In the Introduction, we included now the other references you provided, as well as another one (new text in red). We kept our reference to Hahn & Savin (2014) in the discussion.

-----------------------------------------------------------------------------------------
Q9. Some sections can benefit from weeding out of repetitions. You tend to repeat statements regarding the method a lot. For example:

- Paragraph starting page 13, line 41. This has already been said in the method.
- Sentence Page 14, line 25 'For each leg…
- statement in parenthesis at page 15, line 24.
- Section 4, first two paragraphs.
- Page 26, line 26:
 'In the streamer belt region we combined the DEMT results and the AWSom magnetic model to infer the energy input phi at the coronal base required to maintain stable coronal loops. We found characteristic values in the range…'
Shorten to: 'The energy input phi at the coronal base, required to maintain stable coronal loops, is in the range…'
- Page 26, lines 38-43, between 'Results of the AWSoM….summarized as follows.' can be omitted.

A9. Thanks or the careful read. We have improved the text in those passages, and others similar that we found, with changes highlighted in red.

@@

-----------------------------------------------------------------------------------------
Q10. Minor grammatical improvements, typos, and removal of redundant statements

A10. Thanks very much for this detailed corrections. We believe we addressed them all. Changes highlighted in red.

### 2020-03-11 lloveras_2020_v2.tex
Albert está trabajando en esto, no tocar todavía, gracias.

\begin{abstract}
Observational techniques play an essential role in advancing our understanding of the physical processes governing the solar corona, providing in {particular validation} data for three-dimensional (3D) magnetohydrodynamic (MHD) models of the solar atmosphere. Solar rotational tomography is currently the sole observational technique {able to provide} a 3D quantitative description of the solar corona at a global scale. Based on {extreme ultraviolet} images {of space} borne instruments, {differential emission measure tomography (DEMT)} allows constructing 3D maps of the coronal electron density and temperature at heliocentric heights below $\approx 1.25\,\mrsun$. We carry out a study of the corona {through DEMT} reconstructions {and MHD simulations, using} the latest version of the Alfv\'en Wave Solar Model (AWSoM) developed at the University of Michigan as a component of the Space Weather Modeling Framework. {Two target} rotations were selected from the solar minimum between solar cycles (SCs) 23 and 24 and {the {end of the} declining} phase of SC 24. The 3D {DEMT} and MHD models are analysed and compared in distinct coronal magnetic structures. {We report in quantitative detail on the 3D thermodynamic structure of the inner corona as revealed by the DEMT analysis. We {show the} existence of diverse types of temperature structures in the low and mid latitudes of the {streamer belt. We estimate} the heating energy required to keep them {stable, which we argue may} be accounted for by Alfvén waves. {We also show that results of the latest version of the AWSoM model overall} agree with DEMT {reconstructions within a $\approx 20\%$ accuracy. We highlight some details of the AWSoM model that should} be improved to better reproduce {tomographic} results, in particular concerning the source region of the {fast and slow components of the solar wind.}}
\end{abstract}

>>>>

\begin{abstract}
\edit{We carry out a study of the global three-dimensional (3D) structure of the electron density and temperature of the quiescent inner solar corona ($r<1.25\,\mrsun$) by means of tomographic reconstructions and magnetohydrodynamic simulations. We use differential emission measure tomography (DEMT) and the Alfv\'en Wave Solar Model (AWSoM), in their latest versions. Two target rotations were selected from the solar minimum between solar cycles (SCs) 23 and 24 and the declining phase of SC 24. We report in quantitative detail on the 3D thermodynamic structure of the core and outer layers of the streamer belt, and of the high latitude coronal holes (CH), as revealed by the DEMT analysis. We report on the presence of two types of structures within the streamer belt, loops with temperature decreasing/increasing with height (dubbed down/up loops), as reported first in previous DEMT studies. We also estimate the heating energy flux required at the coronal base to keep these structures stable, found to be or order $10^5\,\erg\,\cminvs\,\s^{-1}$, consistently with previous DEMT and spectroscopic studies. Our findings may explained by Alfvén wave dissipation. We also compare the 3D DEMT and AWSoM results in distinct magnetic structures. We show that the agreement between the products of both techniques is the best so far, with an overall agreement $\lesssim 20\%$, depending on the target rotation and the specific coronal region. Even so, in its current version the ASWsoM model can not reproduce down loops. Also, the model exhibits a latitude variation of the electron density in the source region of the fast and slow components of the solar wind that differs from that of the DEMT results.}
\end{abstract}
